---
title: "DATA607 - HW 3"
author: "Cesar L. Espitia"
date: "February 16, 2017"
output: html_document
---

## Homework Chapter 8

##Homework 3, manipulate the simpsons data.  

```{r simpsons, echo=TRUE}
library(stringr)
snames <- c("Moe Szyslak","Burns, C. Montgomery","Rev. Timothy Lovejoy","Ned Flanders","Simpson, Homer","Dr. Julius Hibbert")

## 3.1 - Generate names First Last
# split vector using strsplit and then apply reversal 
strsplit(snames, split=", ")
#apply over vector reversal and past where split created space
first_last <- sapply(strsplit(snames, split=", "),function(y) {paste(rev(y),collapse=" ")})
print(first_last)

## 3.2
#used str_detect to see if a test is true or false
#specific test
title<-cbind(snames, unlist(str_detect(snames, "Dr.|Rev.")))
print(title)

#unknown test for more generic, requries all to follow convention Alpha Letters + period.
second<-cbind(snames, unlist(str_detect(snames, "[:alpha:]{2,}[.]")))
print(second)

## 3.3
#used str_detect to see if a test is true or false
#specific test
cbind(snames, str_detect(snames," [A-Z]\\."))
```


##Homework 4, describe the expressions and give examples.

(a) [0-9]+\\$
[0-9] describes all digits 0 through 9
"+" says that any of the previous can be matches as many times as possibe
\\$ indicates the dollar symbol at the end

An example of this would be 10000$ but not $100

(b) \\b[a-z]{1,4}\\b
the structure \\b indicates a word wrapped with letters at the start and edge
[a-z] indicates lowercase alpha letters that can be of length 1 through 4

example data, car, at, a.  would not allow I, DOG or others

(c) .*?\\.txt$
"." indicates any item, that is a alpha,digit, or characters
"*" means that the preceding can be matched zero plus
"?" means it is optional but must be matched at most once
\\.txt$ indicates that the .txt must be at the end of the string.  This is different from \\$ which is used to identify a special symbol like ! @ etc.

example would be cesar.txt inclusive of .txt itself

(d) \\d{2}/\\d{2}/\\d{4}
\\d indicates that digits are expected like [:digits:]
{n} indicates the amount expected
/ is an actual forward slash symbol

so this is expecting a date type string
mm/dd/yyyy or dd/mm/yyyy (imperial or metric dates)

(e) <(.+?)>.+?</\\1>
\\1 indicates copy of the prior exact symbol
() is like foil match and says items insides must be matched
".+?" indictes that it can be any characters of any length but must occur only once as a whole in this case

"<html>cesar</html>" would work but "<html>cesar</HTML>" would not

```{r describe, echo=TRUE}
##
samples <- cbind(c("a","a","b","b","c","c","d","d","e","e"),c("10000$","$100","car","DOG","cesar.txt",
          "-.txt","02/15/2017","2/5/2017","<html>cesar</html>","<html>cesar</HTML>"))
df<-cbind(samples,str_detect(samples[,2], "[0-9]+\\$"),str_detect(samples[,2], "\\b[a-z]{1,4}\\b"),
      str_detect(samples[,2], ".*?\\.txt$"),str_detect(samples[,2], "\\d{2}/\\d{2}/\\d{4}"),str_detect(samples[,2], "<(.+?)>.+?</\\1>"))
colnames(df)<-c("3.4","Sample String","a","b","c","d","e")
print(df)
```

please note that the test for c returns TRUE but the sample string to test c also rings true if you use the test from b.  This is becasue b says that the string just has to start or end with an alpha character.  What I am unsure of is why cesar.txt caused a trust if it is clear it is more than 4 letters.  

##Homework 9, manipulate the code.  

This is asking to crack the code, so I will use str_extract_all

```{r hiddencode, echo=TRUE}
code <- c("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")

#tried alpha and too many letters
str_extract_all(code,"[:alpha:]")

#tried lower and garbage 
str_extract_all(code,"[a-z]")

#tried all capital letters and the message shows
str_extract_all(code,"[A-Z]")

#reading back I noticed that the periods parce the text even more.
str_extract_all(code,"[A-Z]|\\.")

```

I am glad I am congratulated for being a super nerd "CONGRATULATIONS YOU ARE A SUPERNERD" #NERD
