---
title: "HW 5"
author: "Cesar L. Espitia"
date: "March 1, 2017"
output: html_document
---

The following is my homework assignment for Week 5 (TIDY/DPLYR).

I am loading the data from MySQL and all the libraries I will use (tidyr, dplyr, knitr, stringr, RMySQL).

```{r setup, echo=TRUE}

library(tidyr)
library(dplyr)
library(stringr)
library(knitr)
library(RMySQL)

con <- dbConnect(RMySQL::MySQL(),
                 dbname = "CUNY607",
                 host = "localhost",
                 port = 3306,
                 user = "root",
                 password = "CUNYDATA")

table_names <- dbListTables(con)
HW5 <- dbReadTable(con, "HW5")

# Fill in missing arilines before doing cleanup
HW5[HW5$Performance == "delayed" & HW5$LosAngeles == "62" &  HW5$Airline == "", "Airline"] <- "ALASKA"
HW5[HW5$Performance == "delayed" & HW5$LosAngeles == "117" &  HW5$Airline == "", "Airline"] <- "AM WEST"

kable(HW5, caption="Table 1. Data from MySQL with empty cells filled in")
```

Cleaning the data with dplyr and tidyr.


```{r cleanup, echo=TRUE}

HW5<-tbl_df(HW5)
HW5a <- HW5 %>% filter(Performance != "")  %>% gather("Destination","Count",4:8)
HW5a <- HW5a[2:5]
HW5a <- HW5a %>% group_by(Airline, Performance, Destination) %>%spread(Performance, Count)
colnames(HW5a) <- str_to_title(colnames(HW5a))
kable(HW5a, caption="Table 2. Airline Data Tidy and Spread for Analysis")

```

Analyze Data

```{r analzye, echo=TRUE}

#prep data for analysis
HW5b <- HW5a %>% mutate(Total = Delayed+`On Time`, PercentDelayed = round(Delayed/Total,2)) 

#Method 1
HW5c <- HW5b %>% group_by(Airline) %>% summarise(AirlineDelay = mean(PercentDelayed,2))
kable(HW5c, caption="Table 3. Airline Total Delay by Taking Mean of Percent Delay of each Destination")

#Method 2
HW5d <- HW5b %>% group_by(Airline) %>% summarise(TotDelay = sum(Delayed),TotOnTime = sum(`On Time`), AirTotal = sum(Total)) %>% mutate(AirDelay = round(TotDelay/AirTotal,2))
kable(HW5d, caption="Table 4. Airline Total Delay by summing all counts then doing a percentage")

```

Note that table 3 and 4 show different values based upon two approaches.  Table 4 would be the correct one for this analysis and show that Alaska Airlines has less delayes total than America West by 4% points.   

Notes:  I had to use stackoverflow to determine why when I used the spread function, I would get NAs with data staggered.  I realized in reviewing comments that I needed to group the data first.  